%-------------------- Research Plan --------------------------
\section{Research Plan, Milestones, and Deliverables}
\textit{This section does not count towards the page limit.}

The dissertation period runs for approximately 11 weeks (June--August 2026).
The core codebase (steering vector extraction, HardConcrete gate training, TruthfulQA evaluation, and all baselines) has already been implemented.
Preliminary results (Section~\ref{sec:preliminary-results}) have revealed an output collapse problem and a length bias confound in the MC evaluation, both of which inform the remaining work plan.

\textbf{WP1: Diagnose and address output collapse (Weeks 1--3).}
Investigate why sparse steering collapses to whitespace on TruthfulQA. Explore scale regularisation, gradient clipping, alternative training objectives, and whether one of the ablated variants (e.g.\ scale-only) avoids the issue. Establish whether the problem is specific to TruthfulQA's training data or fundamental to the technique.
\emph{Dependency: none.}

\textbf{WP2: Expand evaluation suite (Weeks 2--4).}
Finalise the benchmark selection beyond TruthfulQA. Implement evaluation pipelines for candidate tasks (e.g.\ hh-rlhf, JoLA NLU/NLG tasks). Prioritise generative evaluation as the primary success criterion.
\emph{Dependency: none; can run in parallel with WP1.}

\textbf{WP3: Core experiments (Weeks 4--7).}
Run the full comparison across benchmarks: sparse steering versus CAA, ITI, dense, gates-only, scale-only, JoLA, and LoRA. Evaluate along all axes in Section~\ref{sec:evaluation-criteria}: task performance, ablations, parameter efficiency, data efficiency, and robustness across seeds.
\emph{Dependency: WP1 (collapse addressed or understood), WP2 (benchmarks ready).}

\textbf{WP4: Analysis and writing (Weeks 6--11).}
Conduct localisation analysis (gate pattern visualisations, comparison with JoLA/ITI/REAL head selections). Write dissertation chapters. A first complete draft is targeted by Week 9, with two weeks for revision.
\emph{Dependency: WP3 for core results}

\textbf{Meetings.} I suggest weekly 30-minute meetings with Prof.~Titov throughout, with additional ad-hoc check-ins as needed.

\textbf{Resources.} Development hardware: personal laptop.
University cluster access for experiments on larger models once setups are validated.
All datasets and model weights are publicly available via Hugging Face.

\begin{figure}[htbp]
\centering
\begin{ganttchart}[
    y unit title=0.4cm,
    y unit chart=0.5cm,
    vgrid,hgrid,
    x unit=1.4mm,
    time slot format=isodate,
    title/.append style={draw=none, fill=accentblue},
    title label font=\sffamily\bfseries\color{white},
    bar/.append style={draw=none, fill=mainblue},
    bar height=.6,
    bar label font=\small\color{black!50},
    progress label text={}
]{2026-06-01}{2026-08-14}
   \gantttitlecalendar{month=name}\\
   \ganttbar[progress=0]{WP1: Diagnose collapse}{2026-06-01}{2026-06-19} \\
   \ganttbar[progress=0]{WP2: Expand eval suite}{2026-06-08}{2026-06-26} \\
   \ganttbar[progress=0]{WP3: Core experiments}{2026-06-22}{2026-07-17} \\
   \ganttbar[progress=0]{WP4: Analysis \& writing}{2026-06-29}{2026-08-14} \\
   \ganttmilestone{$M_1$: Collapse understood}{2026-06-19} \\
   \ganttmilestone{$M_2$: Benchmarks ready}{2026-06-26} \\
   \ganttmilestone{$M_3$: Core results}{2026-07-17} \\
   \ganttmilestone{$M_4$: First draft}{2026-07-31} \\
   \ganttmilestone{$M_5$: Submission}{2026-08-14}
\end{ganttchart}
\caption{Gantt chart of project activities (June--August 2026).}
\label{fig:gantt}
\end{figure}

%-------------------- Milestones -------------------------
\begin{table}[htbp]
\centering
\begin{tabular}{@{}ccp{8cm}@{}}
\toprule
\textbf{Milestone} & \textbf{Week} & \textbf{Description} \\
\midrule
$M_1$ & 3  & Output collapse diagnosed; mitigation strategy established or negative result documented \\
$M_2$ & 4  & Evaluation suite finalised and pipelines implemented for all benchmarks \\
$M_3$ & 7  & Core experiments complete across all benchmarks and baselines \\
$M_4$ & 9  & First complete dissertation draft \\
$M_5$ & 11 & Final submission \\
\bottomrule
\end{tabular}
\caption{Project milestones.}
\label{tab:milestones}
\end{table}

%-------------------- Deliverables -----------------------
\begin{table}[htbp]
\centering
\begin{tabular}{@{}ccp{8cm}@{}}
\toprule
\textbf{Deliverable} & \textbf{Week} & \textbf{Description} \\
\midrule
$D_1$ & 4  & Open-source codebase with all methods, benchmarks, and evaluation pipelines \\
$D_2$ & 8  & Experimental results: comparison tables, gate visualisations, ablation plots, generative eval outputs \\
$D_3$ & 11 & Completed M.Sc.\ dissertation \\
\bottomrule
\end{tabular}
\caption{Project deliverables.}
\label{tab:deliverables}
\end{table}
